---
title: "DataViz Makeover 2"
description: |
  Explore different perspectives and approaches to design more truthful and enlightening data visualisation
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Data Visualisation

This [interactive data visualisation](https://public.tableau.com/views/prototype1_16468762782120/Dashboard1?:language=en-US&publish=yes&:display_count=n&:origin=viz_share_link) aims to reveal inter- and intra-zonal public bus flows at the planning sub-zone level of January 2022. 


## The Task

1. Critic the data visualisation from its clarity, aesthetics and interactive design.  At least three from each evaluation criterion.  (15 marks)
2. With reference to the critics above, suggest alternative graphical presentation to improve the current design.  Sketch out the proposed design.  The proposed alternative design should include **appropriate interactive techniques**.  Support your design by describing the advantages or which part of the issue(s) your alternative design try to overcome.  (15 marks)
3. Using Tableau, design the proposed data visualisation.  (25 marks)
4. Provide step-by-step description on how the data visualisation was prepared.  (30 marks)
5. Describe three major observations revealed by the data visualisation prepared.
(15 marks)

### Additional design instruction
1. Only data provided should be used, no additional data sets are required.
2. Focus on telling data story using appropriate statistical graphic methods, graphical design principles and interactive techniques, including animation if necessary.


## Data Source
Two data sets are provided for this exercise.  They are:

+ *origin_destination_bus_SZ_202201.csv*, this data set was downloaded from LTA Datamall and URA region, planning area and planning subzone information were added by using GIS Overlay operation.
+ *MP14_SUBZONE_WEB_PL*, this data provides URA Master Plan 2014 Planning Subzone boundary map in ESRI shapefile format.  It was downloaded from data.gov.sg.


## Submission deadline
Submission deadline: **27th March 2022 (Sunday), mid-night 11:59pm**.

## Submission


## Peer Learning

### Quality of Critic

Beside **appropriateness** and **correctness**, a good critic should support with specific examples.

### Quality of Alternative DataViz Makeover Design

DataViz Makeover aims to **improve existing design** as much as possible according to the critics provided.  If possible, students should **avoid introducing a totally new data visualisation**.  You should also **avoid changing the original content of the original design**.  

### Level of Tableau Desktop Proficiency

To demonstrate your Tableau Desktop proficiency, students should avoid using default features and functions as much as possible.  When it is appropriate, you should not hesitate to use advanced functions such as calculated field and parameters.

### Quality of Reproducibility

One of the important learning objective of DataViz Makeover is **to embark your the skill in creating reproducible data visualisation**.  This is achieved by requiring students to to prepare complete and comprehensive step-by-step guide on  how the proposed dataviz makeover is created.  

### Quality of Storytelling

Visual communication plays an important role in a visual analytics process.  No matter how advanced and sophisticated data visualisation techniqiues are used, if we failed to tell a compelling story by using the data visualisation designed, all the hardwork and efforts will be wasted.  


```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("DataViz2.html")
```




