---
title: "Hands-on Exercise 10: Information Dashboard Design: R methods"
subtitle: ""
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
Month: "2020-7-4 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/text.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/macros.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

## Overview

.large[
By the end of this hands-on exercise, you will be able to:

- create bullet chart by using ggplot2,
- create sparklines by using ggplot2  ,
- build industry standard dashboard by using R Shiny. 
]

---
## Importing Microsoft Access database

### The data set
For the purpose of this study, a personal database in Microsoft Access mdb format called *Coffee Chain* will be used.

--
### Importing database into R

In the code chunk below, `odbcConnectAccess()` of [**RODBC**](https://cran.r-project.org/web/packages/RODBC/index.html) package is used used to import a database query table into R. 

```{r eval=FALSE}
library(RODBC)
con <- odbcConnectAccess('data/Coffee Chain.mdb')
coffeechain <- sqlFetch(con, 'CoffeeChain Query')
write_rds(coffeechain, "data/CoffeeChain.rds")
odbcClose(con)
```

Note: Before running the code chunk, you need to change the R system to 32bit version.  This is because the `odbcConnectAccess()` is based on 32bit and not 64bit 

---
## Getting started

.pull-left[
.large[
For the purpose of this hands-on exercise, the following R packages will be used.

```{r}
packages = c('tidyverse', 'lubridate', 
             'ggthemes')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```
]]

--
.pull-right[
.large[
The code chunk below is used to import *CoffeeChain.rds* into R. 


```{r}
coffeechain <- read_rds("data/CoffeeChain.rds")
```

Note: This step is optional if *coffeechain* is already available in R.
]]

---
### Data Preparation

.pull-left[
The code chunk below is used to aggregate Sales and Budgeted Sales at the Product level.

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```
]

---
### Bullet chart in ggplot2 

.pull-left[
The code chunk below is used to plot the bullet charts using ggplot2 functions.

```{r eval=FALSE}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```
]

--
.pull-right[
```{r echo=FALSE, fig.height=8}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```
]

---
## Plotting sparklines using ggplot2

.pull-left[
In this section, you will learn how to plot sparklines by using ggplot2.

Preparing the data

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```
]

.pull-right[
The code chunk below is used to compute the minimum, maximum and end othe the month sales.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))

```

The code chunk below is used to compute the 25 and 75 quantiles.  

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)

```
]

---
### sparklines in ggplot2
The code chunk used.

```{r eval=FALSE, fig.height=3, fig.width=6}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

---
### sparklines in ggplot2
The output

```{r echo=FALSE, fig.height=4, fig.width=9}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, 
              aes(ymin = quart1, 
                  max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), size = 2.5, vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), size = 2.5, vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), size = 2.5, hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), size = 2.5, hjust = 0, nudge_x = 1) +
  expand_limits(x = max(sales_report$Month) + (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 10, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("Hands-on_Ex10.html")
```
