---
title: "Hands-on Exercise 2: Beyond ggplot2 Fundamentals"
subtitle: ""
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "2020-2-15 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/text.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/macros.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.retina=3,
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE)
```

# Content

.large[
- Beyond ggplot2 Themes 

- Beyond ggplot2 Annotation

- Beyond ggplot2 facet 
]

---
# Getting started
## Installing and loading the required libraries

.pull-left[
.large[
- Before we get started, it is important for us to ensure that the required R packages have been installed.  If yes, we will load the R packages.  If they have yet to be installed, we will install the R packages and load them onto R environment.

- The chunk code on the right will do the trick.
]] 

.pull-right[
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('tidyverse', 'ggdist', 'ggridges',
             'patchwork', 'ggthemes', 'hrbrthemes',
             'ggrepel', 'ggforce')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```
]

---
## Importing data

.pull-left[
.large[
- The code chunk below imports *exam_data.csv* into R environment using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package.

- **readr** is one of the tidyverse package.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
exam_data <- read_csv("data/Exam_data.csv")
```
]]

.pull-right[
.large[
- Year end examination grades of a cohort of primary 3 students from a local school.

- There are a total of seven attributes.  Four of them are categorical data type and the other three are in continuous data type.
    - The categorical attributes are: ID, CLASS, GENDER and RACE.
    - The continuous attributes are: MATHS, ENGLISH and SCIENCE.
]]

---
## Beyond ggplot2 Annotation

.pull-left[
One of the challenge in plotting statistical graph is annotation, especially with large number of data points.

```{r eval=FALSE}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
]

.pull-right[

```{r echo=FALSE}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
]

---
### Working with ggrepel

.pull-left[
[**ggrepel**](https://ggrepel.slowkow.com/index.html) is an extension of **ggplot2** package which provides `geoms` for **ggplot2** to repel overlapping text as in our examples on the right. We simply replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html).

```{r eval=FALSE}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
]

```{r echo=FALSE}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

---
## Beyond ggplot2 Themes

.pull-left[
ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html), they are: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, and `theme_void()`.

```{r eval=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +  #<<
  ggtitle("Distribution of Maths scores") 

```

Refer to this [link](https://ggplot2.tidyverse.org/reference/index.html#themes) to learn more about ggplot2 `Themes`,
]

.pull-right[

```{r echo=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 

```
]

---
### Working with ggtheme package

.pull-left[
[**ggthemes**](https://cran.r-project.org/web/packages/ggthemes/index.html) provides ['ggplot2' themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) that replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), 'Stata', 'Excel', and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

```{r eval=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist() #<<
```

It also provides some extra geoms and scales for 'ggplot2'.  Consult [this vignette](https://mran.microsoft.com/snapshot/2016-12-28/web/packages/ggthemes/vignettes/ggthemes.html) to learn more. 
]


.pull-right[
```{r echo=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
]

---
### Working with hrbthems package

.pull-left[
[**hrbrthemes**]() package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used. 

```{r eval=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum() #<<
```

- The second goal centers around productivity for a production workflow. In fact, this “production workflow” is the context for where the elements of hrbrthemes should be used. Consult [this vignette](https://cran.r-project.org/web/packages/hrbrthemes/vignettes/why_hrbrthemes.html) to learn more.
]

.pull-right[
```{r echo=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
]

---
### Working with hrbthems package

.pull-left[
What can we learn from the code chunk below? 

```{r eval=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18, #<<
              base_size = 15, #<<
              grid = "Y") #<<
```

- `axis_title_size` argument is used to increase the font size of the axis title to 18,
- `base_size` argument is used to increase the default axis label to 15, and
- `grid` argument is used to remove the x-axis grid lines.

]

.pull-right[
```{r echo=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```
]

---
## Beyond ggplot2 facet

.pull-left[
In this section, you will learn how to create composite plot by combining multiple graphs.  First, let us create three statistical graphics.  


```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```


```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```
]

.pull-right[

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
]

---
### Creating Composite Graphics

.pull-left[
It is not unusual that multiple graphs are required to tell a compelling visual story.  There are several ggplot2 extensions provide functions to compose figure with multiple graphs.  In this section, I am going to shared with you [**patchwork**](https://patchwork.data-imaginist.com/index.html).  

Patchwork package has a very simple syntax where we can create layouts super easily. Here's the general syntax that combines:
- Two-Column Layout using the Plus Sign +.
- Parenthesis () to create a subplot group.
- Two-Row Layout using the Division Sign `\`
]

---
### Working with patchwork

.center[
```{r fig.width=12, fig.height=6.5}
p1 + p2 / p3
```
]

---
### Working with patchwork

 | will place the plots beside each other, while / will stack them.

.center[
```{r fig.width=12, fig.height=6}
(p1 / p2) | p3
```
]

To learn more about, refer to [Plot Assembly](https://patchwork.data-imaginist.com/articles/guides/assembly.html).
---
### Working with patchword

**patchwork** also provides auto-tagging capabilities, in order to identify subplots in text:

.center[
```{r fig.width=12, fig.height=6}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I') #<<
```
]

---
### Working with patchwork

.center[
```{r fig.width=12, fig.height=6}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
]

---
### Working with patchwork

.pull-left[
Beside providing functions to place plots next to each other based on the provided layout. With [`inset_element()`](https://patchwork.data-imaginist.com/reference/inset_element.html) of **patchwork**, we can place one or several plots or graphic elements freely on top or below another plot. 

```{r eval=FALSE}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```
]

.pull-right[
```{r echo=FALSE}
p12 <- p1|p2
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```
]

---
## Reference

- [A ggplot2 Tutorial for Beautiful Plotting in R](https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/#text)


```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("Hands-on_Ex02.html")
```
