---
title: 'Hands-On Exercise 5: Building Parallel Coordinates Plot with R'
subtitle: ""
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems"
institute: "School of Information Systems,<br/>Singapore Management Uniiversity"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

# 1.0 Overview

Parallel coordinates plot is a data visualisation specially designed for visualising and analysing multivariate, numerical data.  It is ideal for comparing multiple variables together and seeing the relationships between them. For example, the variables contribute to Happiness Index. 

In this hands-on exercise, you will learn how to plot parallel coordinates plot programmatically by using appropriate R packages.

# 2.0 Installing and Launching R Packages

For this exercise, the **GGally**, **plotly R**, **parcoords** and **tidyverse** packages will be used.  

The code chunks below are used to install and load the packages in R.

```{r}
packages = c('GGally', 'plotly', 'parcoords', 'parallelPlot', 'tidyverse')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# 3.0  Data Preparation

In this hands-on exercise, the World Happinees 2018 (http://worldhappiness.report/ed/2018/) data will be used.  The data set is download at https://s3.amazonaws.com/happiness-report/2018/WHR2018Chapter2OnlineData.xls.  The original data set is in Microsoft Excel format.  It has been extracted and saved in csv file called **WHData-2018.csv**. 


# 3.1 Importing World Happiness Data Set 

In this section, you will learn how to visualise and analyse multiple continuous variables using correlation matrix, parallel coordinate plot and heatmap.  World Happiness Report 2018 dataset (WHData-2018.csv) will be used in this section.

First, important the data into R by using the code below.

```{r, echo=TRUE, message=FALSE}
wh <- read_csv("data/WHData-2018.csv")
```

# 4.0 Working with *ggparcoord()* of **GGally** package

## 4.1 Plotting a simple parallel coordinates

The code chunk below is used to plot a basic parallel coordinates by using [*ggparcoord()*](https://ggobi.github.io/ggally/reference/ggparcoord.html) of **GGally** package.

```{r fig.width=13, fig.height=6}
ggparcoord(data = wh, 
           columns = c(7:12))
```

## 4.2 Plotting a parallel coordinates with boxplot

- Underlay univariate boxplots, add title, use uniminmax scaling

```{r, echo=TRUE, fig.width=13, fig.height=5}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coord. Plot of World Happines Attributes")
```

## 4.3 Parallel coordinates with facet

- Underlay univariate boxplots, add title, use uniminmax scaling

```{r, echo=TRUE, fig.width=14, fig.height=6}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax", 
           boxplot = TRUE, 
           title = "Parallel Coord. Plot of World Happines Attributes") +
  facet_wrap(~ Region)
```

# 5.0 Plotting Interactive Parallel Coordinates: parcoords methods

## 5.1 Getting to know **parcoords** package

**parcoords** is a htmlwidget for d3 parallel coordinates plot.  It gives R users the very well designed and interactive parallel-coordinates chart for d3 with the infrastructure, flexibility, and robustness of htmlwidgets.

## 5.2 The basic plot

The code below plot a parallel coordinates plot using the basic *parcoords()* syntax.

```{r, fig.width=10, fig.height=5}
parcoords(wh[,7:12])
```

In fact, the basic syntax can only plot a static parallel coordinates plot.

## 5.3 Plotting an Interactive Parallel Coordinates Plot

In the code chunk below, two interactivity arguments are used, they are:

+ `reorderable` argument enables reordering of axes when it is set to TRUE or T, and
+ `brushMode` argument provides highlighting with brushing. Three options are provided, they are: "1D-axes", "1D-axes-multi", or "2D-strums". 

```{r, fig.width=10, fig.height=6}
parcoords(
 wh[,7:12],
 rownames = FALSE,
 reorderable = T,
 brushMode = '1D-axes',
 width = 700,
 height = 400)
```

Other things to learn from the code chunk above.

+ By setting the `rownames` argument to FALSE will exclude the default rawnames column.
+ By using `width` argument to avoid horizontal scrolling.
+ By using `height` argument to change the height of the parallel coordinates plot.

## Changing aethetic properties

In the code chunk below, two arguments are used, they are:

+ `alpha` to set the opacity of the polylines.  The value range from 0 (completely transparent) to 1 (no opacity).
+ `alphaOnBrushed` to set opacity from 0 to 1 when brushed (default to 0).

```{r, fig.width=10, fig.height=6}
parcoords(
 wh[,7:12],
 rownames = FALSE,
 alpha = 0.3,
 alphaOnBrushed = 0.2,
 reorderable = T,
 brushMode = '1D-axes',
 width = 700,
 height = 400)
```

# 6.0 Plotting Interactive Parallel Coordinates Plot: parallelPlot methods

[**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/) is an R package specially designed to plot a parallel coordinates plot by using 'htmlwidgets' package and [d3.js](https://d3js.org/).  In this section, you will learn how to use functions provided in parallelPlot package to build interactive parallel coordinates plot.

## 6.1 The basic plot

The code chunk below plot an interactive parallel coordinates plot by using *parallelPlot()*.

```{r fig.align='center', fig.width=12, fig.height=4}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh)
```

Notice that some of the axis labels are too long.  You will learn how to overcome this problem in the next step.  

### 6.1.1 Rotate axis label

In the code chunk below, `rotateTitle` argument is used to avoid overlapping axis labels.

```{r fig.align='center', fig.width=12, fig.height=4}
parallelPlot(wh,
             rotateTitle = TRUE)
```

One of the useful interactive feature of parallelPlot is we can click on a variable of interest, for example Happiness score, the monotonous blue colour (default) will change a blues with different intensity colour scheme will be used. 

### 6.1.2 Changing the colour scheme

We can change the default blue colour scheme by using `continousCS` argument as shown in the code chunl below.

```{r fig.align='center', fig.width=12, fig.height=4}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```





## 6.2 Parallel coordinates plot with hitogram

In the code chunk below, `histoVisibility` argument is used to plot histogram along the axis of each variables.  

```{r fig.align='center', fig.width=12, fig.height=4}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```





# References

- [*ggparcoord()*](http://ggobi.github.io/ggally/reference/ggparcoord.html) of [**GGally**](http://ggobi.github.io/ggally/index.html) package

- [**parcoords** user guide](https://cran.r-project.org/web/packages/parcoords/parcoords.pdf)

- [**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/parallelPlot.pdf)
