---
title: "Take-home Exercise 1"
description: |
  Creating data visualisation beyond default
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The Task

In this take-home exercise, you are required to apply the skills you had learned in Lesson 1 and Hands-on Exercise 1 to reveal the demographic of the [city of Engagement, Ohio USA](https://vast-challenge.github.io/2022/) by using appropriate static statistical graphics methods.  The data should be processed by using appropriate **tidyverse** family of packages and the statistical graphics must be prepared using **ggplot2** and its extensions.

### Survival Tip

Start with *Participant*s data file in the Attributes folder.

# Submission Instructions 

This is an individual assignment. You are required to work on the take-home exercises and prepare submission individually. 

- The take-home exercise submission must be written by using **distill for R Markdown**.  It can be in either [distill article](https://distill.pub/guide/) or [distill blog ](https://rstudio.github.io/distill/blog.html) format. You are required to publish the exercise on [Netlify](https://www.netlify.com/) and provide the link on the take-home exercise submission page on elearn. 
- You are also required to push the take-home exercise folder and data onto your individual Github repository and provide the link on the take-home exercise submission page on eLearn.
 
### Submission date 
 
Your completed take-home exercise is due on **24th April 2022, by 11:59pm evening**.

### Learning from the senior

- [Singapore’s Demographic Composition Across Time](https://rpubs.com/deniseadele/584423)


# Peer Learning



```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/take-home.html")
```



















