---
title: "Take-home Exercise 2"
description: |
  Creating data visualisation beyond default
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The Task

In this take-home exercise, you are required to apply appropriate interactivity and animation methods you had learned in last week lesson to design an age-sex pyramid based data visualisation to show the changes of demographic structure of Singapore by age cohort and gender between 2000-2020 at planning area level.  

For this task, the data sets entitle *Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2000-2010* and *Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020* should be used.  These data sets are available at [Department of Statistics](https://www.singstat.gov.sg/) home page.   

### Word of advice

Students are strongly encouraged to review the submission done by your peers for [Take-home Exercise 1](https://isss608-ay2021-22t2.netlify.app/th_ex/th_ex01) to either improve or correct some of the mistake you have done in your previous design before attempting this take-home exercise.  This is to ensure that you are not going to repeat the same mistake again.


# Submission Instructions 

This is an individual assignment. You are required to work on the take-home exercises and prepare submission individually. 

- The take-home exercise submission must be written by using **distill for R Markdown**.  It can be in either [distill article](https://distill.pub/guide/) or [distill blog ](https://rstudio.github.io/distill/blog.html) format. You are required to publish the exercise on [Netlify](https://www.netlify.com/) and provide the link on the take-home exercise submission page on elearn. 
- You are also required to push the take-home exercise folder and data onto your individual Github repository and provide the link on the take-home exercise submission page on eLearn.
 
### Submission date 
 
Your completed take-home exercise is due on **4th February 2022, by 11:59pm evening**.

# Peer Learning

+ [ANG BI LIAN](https://bilian-mitb2021.netlify.app/th_ex/th_ex02/take-home_ex02)
+ [ANNIE CHAN WA LAM](https://anniecwl.netlify.app/takehomeex2/takehome-ex2)
+ [CLEMENT ONG SHI QI](https://isss608cosq.netlify.app/th_ex/th_ex02/th_ex02)
+ [FEI JIAHUI](https://fjhva.netlify.app/th_ex/th_ex02/th_ex02)
+ [FOO LI JING FARAH](https://farahfoo.netlify.app/th_ex/th_ex03/th_ex03) 
+ [GONG SHUFEN](https://va2022.netlify.app/th_ex/th_ex02/th_ex02_distill)
+ [HE QIYUN](tps://qyva.netlify.app/th_ex/th_ex02/take-home-exercise02)
+ JIANG MINGJUN
+ [KARISHMA YADAV](https://karishma-smu.netlify.app/th_exercise/th_ex02/takehome_ex02)
+ [KWONG MEI LIN](https://kml-va.netlify.app/th_ex/th_ex02/th_ex02)
+ [LIN SHUYAN](https://lins-92-isss608.netlify.app/th_ex/th_ex02/th_ex02)
+ [LIU CHENTAO](https://amberjourney.netlify.app/th_ex/th_ex2/take-home-exercise_2)
+ [MAK HAN REN](https://makhanren.netlify.app/th_ex/th_ex02/th_ex02)
+ [MIN XIAOQI](https://mxqvisualanalytics.netlify.app/th_ex/take_home_ex02/take_home_ex_2)
+ [NG JIANYUAN CAINE](https://ngcaine.netlify.app/takehomeexercise/ex02/takehomeexercise02)
+ [REYNARD LAM KUAN TING](https://reynard.netlify.app/th_ex/th_ex02/th_ex02)
+ [SEAH XING XIAN](https://sxingxian-va.netlify.app/th_ex/th_ex02/th_ex02)
+ [SEAN SAMUEL PRAJS](https://myanalyticsartgallery2.netlify.app/takehomeex/2/thex2)
+ [SIA HENG GUANG](https://hengguang-608.netlify.app/takehome_exercise/ex002/ex002)
+ [SUN SHENGMEI](https://isss608visualanalyticssun.netlify.app/th_ex/th_ex02/take-home-ex02)
+ [TAN HUI QI RANICE](https://isss608-ranice.netlify.app/th_ex/th_ex02/th_ex02)
+ [TAN JIT KAI](https://tjkva2022.netlify.app/th_ex/th_ex02/th_ex02)
+ [TAN MAY HUA MELISSA  MELISSA XU](https://mviz.netlify.app/takehome_ex02/takehome_ex02)
+ [TAN YAN RU](https://yanrutanfirstsite.netlify.app/th_ex/take-home%20ex02/take-home-ex02)
+ [TING PEILING](https://eloquent-va.netlify.app/th_ex/th_ex02/th_ex02_age_sex_pyramid_animate)
+ [TING YEN-HSIN](https://yhting2021.netlify.app/th_ex/th_ex03/th_ex03)
+ [YANG NENGJIE BENNIE](https://isss608bennie.netlify.app/takehomeex/thex2/thex2)
+ [ZHANG JIEYUAN](https://jyvisualanalytics.netlify.app/takehomeex/takehomeex2/ex2)
+ [ZHANG YI](https://isss608.netlify.app/posts/2022-02-04-take-home-exercise-2/)
+ [ZHOU YIOU]()   

```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/take-home.html")
```



















