---
title: "Take-home Exercise 3"
description: |
  Putting Visual Analytics into Practical Use
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The Task

In this take-home exercise, you are required to apply appropriate data visualisation techniques you had learned in **Lesson 4** or **Lesson 5** to accomplish either one of the task given below:  

1.  Create a data visualisation showing average rating and proportion of cocoa percent (% chocolate) greater than or equal to 70% by top 15 company location.  For the purpose of this task, *chocolate.csv* should be used.

2. Create a data visualisation to provide an unbiased comparison of order returned by state.  For the purpose of this task, *Superstore-2021.xls* should be used.

3. Create a data visualisation to segment kid drinks and other by nutrition indicators. For the purpose of this task, *starbucks_drink.csv* should be used.

Note: 

+ All the data files are available in the **Take-home Exercise 3** folder of course eLearn.
+ For Task 1 or Task 2, refer to **Lesson 4: Fundamental of Visual Analytics** and for Task 3 refer to **Lesson 5: Visual Multivariate Analysis**.

# Submission Instructions 

This is an individual assignment. You are required to work on the take-home exercises and prepare submission individually. 

- The take-home exercise submission must be written by using **distill for R Markdown**.  It can be in either [distill article](https://distill.pub/guide/) or [distill blog ](https://rstudio.github.io/distill/blog.html) format. You are required to publish the exercise on [Netlify](https://www.netlify.com/) and provide the link on the take-home exercise submission page on elearn. 
- You are also required to push the take-home exercise folder and data onto your individual Github repository and provide the link on the take-home exercise submission page on eLearn.
 
### Submission date 
 
Your completed take-home exercise is due on **20th February 2022, by 11:59pm evening**.

# Peer Learning

+ [ANG BI LIAN](https://bilian-mitb2021.netlify.app/th_ex/th_ex03/take-home_ex03)
+ [ANNIE CHAN WA LAM](https://anniecwl.netlify.app/takehomeex3/takehome-ex3)
+ [CLEMENT ONG SHI QI](https://isss608cosq.netlify.app/th_ex/th_ex03/th_ex03)
+ [FEI JIAHUI](https://fjhva.netlify.app/th_ex/th_ex03/th_ex03)
+ [FOO LI JING FARAH](https://farahfoo.netlify.app/th_ex/th_ex03/th_ex03) 
+ [GONG SHUFEN](https://va2022.netlify.app/th_ex/th_ex03/th_ex03_distill)
+ [HE QIYUN](https://qyva.netlify.app/th_ex/th_ex03/exercise3)
+ [JIANG MINGJUN](https://isss608mj.netlify.app/assigment/takehome03/takehome3.html)
+ [KARISHMA YADAV](https://karishma-smu.netlify.app/th_exercise/th_ex03/takehome_ex03)
+ [KWONG MEI LIN](https://github.com/kwongml/ISSS608_VA/tree/master/TH_Ex/TH_Ex03)
+ [LIN SHUYAN](https://lins-92-isss608.netlify.app/th_ex/th_ex03/th_ex03)
+ [LIU CHENTAO](https://amberjourney.netlify.app/th_ex/th_ex3/take-home-exercise_3)
+ [MAK HAN REN](https://makhanren.netlify.app/th_ex/th_ex03/th_ex03)
+ [MIN XIAOQI](https://mxqvisualanalytics.netlify.app/th_ex/take_home_ex03/take_home_ex_03)
+ [NG JIANYUAN CAINE](https://ngcaine.netlify.app/takehomeexercise/ex03/takehomeexercise03)
+ [REYNARD LAM KUAN TING](https://reynard.netlify.app/th_ex/th_ex03/th_ex03)
+ [SEAH XING XIAN](https://sxingxian-va.netlify.app/th_ex/th_ex03/)
+ [SEAN SAMUEL PRAJS](https://myanalyticsartgallery2.netlify.app/takehomeex/3/thex3)
+ [SIA HENG GUANG](https://hengguang-608.netlify.app/takehome_exercise/ex003/ex003)
+ [SUN SHENGMEI](https://isss608visualanalyticssun.netlify.app/th_ex/th_ex03/take_home_ex03)
+ [TAN HUI QI RANICE](https://isss608-ranice.netlify.app/th_ex/th_ex03/th_ex03)
+ [TAN JIT KAI](https://tjkva2022.netlify.app/th_ex/th_ex03/th_ex03)
+ [TAN MAY HUA MELISSA  MELISSA XU](https://mviz.netlify.app/takehome_ex03/takehome_ex03)
+ [TAN YAN RU](https://yanrutanfirstsite.netlify.app/th_ex/take-home%20ex03/take_home_ex03)
+ [TING PEILING](https://eloquent-va.netlify.app/th_ex/th_ex03/th_ex03_cocoa)
+ [TING YEN-HSIN](https://yhting2021.netlify.app/th_ex/th_ex04/th_ex04-1)
+ [YANG NENGJIE BENNIE](https://isss608bennie.netlify.app/takehomeex/thex3/thex3)
+ [ZHANG JIEYUAN](https://jyvisualanalytics.netlify.app/takehomeex/takehomeex3/ex3)
+ [ZHANG YI](https://isss608.netlify.app/posts/2022-02-15-take-home-exercise-3/)
+ [ZHOU YIOU](https://isss608-2022-yiou.netlify.app/th_ex/takehome_ex03/th3)

   

```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/take-home.html")
```



















