---
title: "Take-home Exercise 2"
description: |
  Creating data visualisation beyond default
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The Task

In this take-home exercise, you are required to:

- select one of the Take-home Exercise 1 prepared by your classmate, 
- critic the submission in terms of clarity and aesthetics, and 
- remake the original design by using the data visualisation principles and best practice you had learned in Lesson 1 and 2.

# Submission Instructions 

This is an individual assignment. You are required to work on the take-home exercises and prepare submission individually. 

- The take-home exercise submission must be written by using **distill for R Markdown**.  It can be in either [distill article](https://distill.pub/guide/) or [distill blog ](https://rstudio.github.io/distill/blog.html) format. You are required to publish the exercise on [Netlify](https://www.netlify.com/) and provide the link on the take-home exercise submission page on elearn. 
- You are also required to push the take-home exercise folder and data onto your individual Github repository and provide the link on the take-home exercise submission page on eLearn.
 
### Submission date 
 
Your completed take-home exercise is due on **1st May 2022, by 11:59pm evening**.

# Peer Learning

- [ALOYSIUS TENG]()
- [ANTONIUS HANDY]()
- [CHE XUAN]()
- [CHU YI-NING]()
- [CUI  JIARUI]()
- [DAVID KWOK WEI XIANG]()
- [DING YANMU]()
- [HERANSHAN SO SUBRAMANIAM]()
- [HUANG ANNI]()
- [HUANG YAPING]()
- [HULWANA BINTE SAIFULZAMAN]()
- [KIM BOMIN]()
- [LAM KUO JIAN JEREMIAH]()
- [LEE XIAO QI]()
- [LI HONGYI]()
- [LI HUAN]()
- [LI MINQI]()
- [LIU ZHENGLIN]()
- [LONG NU]() 
- [NI JUNQIU]()
- [ONG ZHI RONG JORDAN]()
- [RAKENDU RAMESH]()
- [RAO NINGZHEN]()
- [RAUNAK KAPUR]()
- [RAVEENA CHAKRAPANI]()
- [SHACHI ANIRUDHA RAODEO]()
- [SUI TIANYUE]()
- [TAY CHENG WEI CLARENCE]()
- [TSENG CHIA-YI]()
- [WANG TIANQI]()
- [YEO KIM SIANG]()
- [YU DI]()
   

```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/take-home.html")
```



















