---
title: "Take-home Exercise 2"
description: |
  Creating data visualisation beyond default
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The Task

In this take-home exercise, you are required to:

- select one of the Take-home Exercise 1 prepared by your classmate, 
- critic the submission in terms of clarity and aesthetics, and 
- remake the original design by using the data visualisation principles and best practice you had learned in Lesson 1 and 2.

# Submission Instructions 

This is an individual assignment. You are required to work on the take-home exercises and prepare submission individually. 

- The take-home exercise submission must be written by using **distill for R Markdown**.  It can be in either [distill article](https://distill.pub/guide/) or [distill blog ](https://rstudio.github.io/distill/blog.html) format. You are required to publish the exercise on [Netlify](https://www.netlify.com/) and provide the link on the take-home exercise submission page on elearn. 
- You are also required to push the take-home exercise folder and data onto your individual Github repository and provide the link on the take-home exercise submission page on eLearn.
 
### Submission date 
 
Your completed take-home exercise is due on **1st May 2022, by 11:59pm evening**.

# Peer Learning

- [ALOYSIUS TENG](https://visual-analytics-course.netlify.app/th_ex/th_ex2.html)
- [ANTONIUS HANDY](https://isss608-hhhandy.netlify.app/th_ex/th_ex2.html)
- [CHE XUAN](https://r4dsa4va.netlify.app/th_ex/th_ex2.html)
- [CHU YI-NING](https://yn-isss608.netlify.app/take-home_ex/th_ex2.html)
- [CUI  JIARUI](https://jiarui-cui-isss608.netlify.app/th_ex/th_ex02)
- [DAVID KWOK WEI XIANG](https://dkwx1234.netlify.app/th_ex/takehomeex2.html)
- [DING YANMU](https://dingyanmu.netlify.app/th_ex/th_ex2.html)
- [HERANSHAN SO SUBRAMANIAM](https://heranshan-isss608.netlify.app/th_ex/th_ex2.html)
- [HUANG ANNI](https://huanganni.netlify.app/th_ex/th_ex2.html)
- [HUANG YAPING](https://vast-challenge-2022-th-ex-1.netlify.app/th_ex/th_ex2.html)
- [HULWANA BINTE SAIFULZAMAN]()
- [KIM BOMIN](https://bomink.netlify.app/posts/2022-04-25-hw2/)
- [LAM KUO JIAN JEREMIAH](https://jeremiahlam.netlify.app/th_ex/th_ex2.html)
- [LEE XIAO QI](https://tiffanylxq.netlify.app/takehomeex/th_ex2)
- [LI HONGYI](https://jessielee.netlify.app/th_ex/th_ex2.html)
- [LI HUAN](https://katherine-isss608.netlify.app/th_ex/th_ex2.html)
- [LI MINQI](https://isss608-visualization-liminqi.netlify.app/th_ex/th_ex2)
- [LIU ZHENGLIN](https://zhenglinliuvisual.netlify.app/th_ex/th_ex2.html)
- [LONG NU](https://leslielong-da.netlify.app/posts/2022-04-25-takehome2/) 
- [NI JUNQIU](https://isss-608-junqiuni.netlify.app/takehome_ex/takehome_ex02.html)
- [ONG ZHI RONG JORDAN](https://isss608-jordan-va.netlify.app/th_ex2/th_ex2.html)
- [RAKENDU RAMESH](https://rakendu-mitb2021.netlify.app/th_ex/th_ex02.html)
- [RAO NINGZHEN]()
- [RAUNAK KAPUR](https://isss608-raunakkapur.netlify.app/th_ex/the2)
- [RAVEENA CHAKRAPANI](https://raveenaclr.netlify.app/th_ex/th_ex2.html)
- [SHACHI ANIRUDHA RAODEO](https://visualanalytics.netlify.app/th_ex/th_ex2.html)
- [SUI TIANYUE](https://isss608totallynew.netlify.app/take_home_exercise/ex_2.html)
- [TAY CHENG WEI CLARENCE](https://clarencetay.netlify.app/th_ex/th_ex2.html)
- [TSENG CHIA-YI](https://mitb-va-joycetseng.netlify.app/posts/2022-04-26-takehome2/)
- [WANG TIANQI](https://wtqmitbva.netlify.app/takehome2)
- [YEO KIM SIANG]()
- [YU DI](https://yudiva.netlify.app/exe2)
   

```{r echo=FALSE, eval=FALSE}
library(pagedown)
pagedown::chrome_print("_site/take-home.html")
```



















